---
title: "Data Import"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

let's import a dataset.

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv")
```

look at my data!
```{r}
names(litters_df)
```

update the names in `litters_df` aka clean any spaces, extra special characters, apply uppercase and lowercase formatting.
note - janitor::clean_names means you are taking the clean_names function from the janitor package without having to load the whole package.
```{r}
litters_df = 
  janitor::clean_names(litters_df)
```

look at my data again! 
note - if you want to look at the data set in a separate window once, type view(dataset name) in console pane below.
```{r}
litters_df
```


##Fix the missingness

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv", na = c("NA", ".", " "))
```

let's import Pups dataset.

```{r}
pups_df = 
  read_csv("data_import_examples/FAS_pups.csv",
  skip = 3,
  na = c("NA", ".", " ")
  )


pups_df = 
  janitor::clean_names(pups_df)
```

## Okay what about excel

CSVs are really great but sometimes you get an excel file.
note - you need to load readxl package for this.

```{r}
mlb_df = 
  read_excel("data_import_examples/mlb11.xlsx")
```

import one part of LotR word counts.
```{r}
fotr_df = 
  read_excel("data_import_examples/LotR_Words.xlsx", range = "B3:D6")
```

## SAS???

let's look at another dataset from SAS.
note - you need to load haven package for this.

import the PULSE data.

```{r}
pulse_df = 
  read_sas("data_import_examples/public_pulse_data.sas7bdat")

pulse_df = 
  janitor::clean_names(pulse_df)
```


## why does my professor hate read.csv so much??

it does not print well so always use read_csv
```{r, eval = FALSE}
litters_df_base = 
  read.csv("data_import_examples/FAS_litters.csv")
```

