---
title: "Data Import"
output: github_document
---

```{r}
library(tidyverse)
```

let's import a dataset.

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv")
```

look at my data!
```{r}
names(litters_df)
```

update the names in `litters_df` aka clean any spaces, extra special characters, apply uppercase and lowercase formatting.
note - janitor::clean_names means you are taking the clean_names function from the janitor package without having to load the whole package.
```{r}
litters_df = 
  janitor::clean_names(litters_df)
```

look at my data again! 
note - if you want to look at the data set in a separate window once, type view(dataset name) in console pane below.
```{r}
litters_df
```


##Fix the missingness

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv", na = c("NA", ".", " "))
```

let's import Pups dataset.

```{r}
pups_df = 
  read_csv("data_import_examples/FAS_pups.csv",
  skip = 3,
  na = c("NA", ".", " ")
  )


pups_df = 
  janitor::clean_names(pups_df)
```
